---
title: "Rodrigues, 2020 (Exercises 1 & 2)"
author: "Amanda Moeller"
date: "2020.03.16"
output: html_notebook
---



## **Exercise 1:**

### Load the LaborSupply dataset from the Ecdat package and answer the following questions:
```{r}
install.packages("Ecdat")
library(Ecdat)
devtools::install_github("tidyverse/tidyr")
library(tidyr)


data(LaborSupply)
head(LaborSupply) 
str(LaborSupply)
```

### Compute the average annual hours worked by year (plus standard deviation)
``` {r}
# ("lnhr" = log of annual hours worked)

# mean hrs worked by year:
mean_hoursBYyear <- aggregate(LaborSupply$lnhr, by = list(year = LaborSupply$year), mean)
mean_hoursBYyear

# SD hours worked by year:
SD_hoursBYyear <- aggregate(LaborSupply$lnhr, by = list(YEAR = LaborSupply$year), sd)
SD_hoursBYyear

# 4.1 (Rodrigues, 2018)
          
```

### What age group worked the most hours in the year 1982?
``` {r}
LaborSupply1 <- LaborSupply %>% filter(year==1982) %>% 
  group_by(age) %>% 
  mutate(WorkHrsByAge=sum(lnhr))%>%
  ungroup() %>%
  filter(WorkHrsByAge==max(WorkHrsByAge))
LaborSupply

# The 34-year old age group worked the greatest amount of hours in 1982, with a  mean of 270.4
```

### Create a variable, n_years that equals the number of years an individual stays in the panel. Is the panel balanced?
``` {r}
# Observations range from 1979 to 1988
# So number of years an individual stays can range from 1-10
LaborSupply %>% group_by(id) %>% mutate(n_years=n())

# It looks like all stay in for 10 years?
# That's probably incorrect. I'll return to this later.
```

### Which are the individuals that do not have any kids during the whole period? Create a variable, no_kids, that flags these individuals (1 = no kids, 0 = kids)
``` {r}
LaborSupply3 <- LaborSupply %>% group_by(id) %>% 
  mutate(totalkids=max(kids)) %>% mutate(no_kids=if_else(totalkids == 0, 1, 0))
LaborSupply3

# column "no_kids" shows 0 for individuals with kids at some point throughout the study and 1 for individuals with no reported kids throughout the study
```

### Using the no_kids variable from before compute the average wage, standard deviation and number of observations in each group for the year 1980 (no kids group vs kids group).
``` {r}
LaborSupply4 <- LaborSupply3 %>% group_by(no_kids) %>% filter(year==1980) %>% summarise(mean(lnwg), sd(lnwg), n())

LaborSupply4

# For individuals with kids, the average wage was 2.59 (sd=0.39) in 1980.
# For individuals without kids, the average wage was 2.68 (sd = 0.42) in 1980.

```
## *Exercise 2:*

### What does the following code do? Copy and paste it in an R interpreter to find out!
```{r}
LaborSupply %>%
  group_by(id) %>%
  mutate_at(vars(starts_with("l")), funs(lag, lead))

# It looks like this code groups by participant and gives us the lag and lead for all variables that begin with the letter "l" (lnhr and lnwg)
```


### mutate_at() is a scoped version of mutate() which allows you to specify a number of columns and functions in one go. This also exists for summarise().

### Using summarise_at(), compute the mean, standard deviation and number of individuals of lnhr and lnwg for each individual.
```{r}
LaborSupply %>%
  group_by(id) %>%
  mutate_at(vars(starts_with("l")), funs(mean, sd, n()))
# lnwg_mean = 1.95
# lnwg_sd = 0.07
# lnwg_n = 10

# lnhr_mean = 7.65
# lnhr_sd = 0.14
# lnhr_n = 10

```



### Create the lagged logarithm of hours worked and wages. Remember that this is a panel.
``` {r}
# I'm not super knowledgeable about what this means, but I'm just going to wing it

LaborSupply %>% 
  group_by(id) %>%
  mutate(lag_hours = lag(lnhr)) %>%
  mutate(lag_lnwg = lag(lnwg)) 

# ... Did I do it?
```


